LaMachine: A meta-distribution for NLP software

### 2019-03-29 docs/lamachine_clarin2018.tex
minor edits
Various cropping of images, shortening and merging phrases, to get everything within the page limit whilst still offering better readability in the images

%\setlength\titlebox{5cm}
\setlist{noitemsep}

@@

\section{ Included Software}

@@

     Nijmegen\footnote{Links: \url{https://languagemachines.github.io/timbl}, \url{https://languagemachines.github.io/mbt},

@@

 \item by the University of Groningen
     \item \textbf{Alpino}\footnote{\url{http://www.let.rug.nl/vannoord/alp/Alpino/}} - A dependency parser and tagger for Dutch.
 \item by the Vrije Universiteit Amsterdam
     \item \textbf{KafNafParserPy}\footnote{\url{https://github.com/cltl/KafNafParserPy}} - A python module to parse NAF
         files, another format for linguistic annotation.
     \item \textbf{NafFoLiApy}\footnote{\url{https://github.com/cltl/NafFoLiApy}} - Converters between FoLiA and NAF.
 \item by Utrecht University
     \item  \textbf{T-scan}\footnote{\url{https://github.com/proycon/tscan}} - T-scan is a Dutch text analytics tool for readability prediction (initially developed at TiCC, Tilburg University, and in collaboration with Radboud University, Nijmegen)
 \item by the Meertens Instituut
     \item \textbf{Python Course for the Humanities}\footnote{\url{http://www.karsdorp.io/python-course/}} - Interactive

@@

\begin{itemize}
    \item \textbf{Python}: Numpy, Scipy, Matplotlib, Scikit-Learn, IPython, Jupyter, ...
    \begin{itemize}
        \item \textbf{Jupyter Lab} The successor of the popular Jupyter Notebooks, offers notebooks, a web-based IDE, terminals. An ideal entry point to get started with LaMachine and all it contains!
        \item \textbf{PyTorch} - Deep-learning library for Python
        \item \textbf{NLTK} - Natural Language Toolkit for Python
        \item \textbf{Spacy} - Industrial-Strength NLP in Python
    \end{itemize}
    \item \textbf{R}
    \item \textbf{Java}
    \begin{itemize}
        \item \textbf{NextFlow} - A system and language for writing parallel and scalable pipelines in a portable manner.
        \item \textbf{Stanford CoreNLP} - Various types of linguistic enrichment
    \end{itemize}
    \item \textbf{Tesseract} - Open Source Optical Character Recognition (OCR)
    \item \textbf{Tensorflow} - Open-source machine learning framework
    \item \textbf{Kaldi} - Speech Recognition Framework (ASR)
    \item \textbf{Moses} - Statistical Machine Translation system
\end{itemize}

@@

\begin{figure}[htb] \begin{center}
\includegraphics[width=110.0mm]{screenshot_bootstrap.jpg}
\end{center}
\caption{\footnotesize{A screenshot of the bootstrap procedure}}
\label{fig:bootstrap}
\end{figure}

@@

\begin{figure}[htb] \begin{center}
\includegraphics[width=135.0mm]{screenshot_venv_activate.jpg}
\end{center}
\caption{\footnotesize{Activation of a local LaMachine environment on the command-line}}
\label{fig:venv}
\end{figure}

@@

\begin{figure}[htb] \begin{center}
\includegraphics[width=135.0mm]{screenshot_portal.jpg}
\end{center}
\caption{\footnotesize{A screenshot of the LaMachine portal, powered by Labirinto:
\url{https://github.com/proycon/labirinto}}}
\label{fig:portal}
\end{figure}

@@

\begin{figure}[htb] \begin{center}
\includegraphics[width=135.0mm]{screenshot_lab.jpg}
\end{center}
\caption{\footnotesize{A screenshot of the Jupyter Lab Environment}}
\label{fig:lab}
\end{figure}

@@

There are also CLARIN initiatives that provide not just a switchboard, but a more elaborate high-level virtual research
environment (VRE) with graphical interfaces to accommodate the researcher and allow execution and visualisation of
various NLP pipeline. The Dutch CLARIAH WP3 project\footnote{\url{https://github.com/meertensinstituut/clariah-wp3-vre}}
is such an initiative. LaMachine is currently developed in
close collaboration with the Dutch CLARIAH WP3 VRE project, in which we act as a technology supplier.

@@

LaMachine is most often used by a researcher or data scientist who installs LaMachine on his own computer (locally, in a
local environment or as a Docker container or VM). Here we discuss two case studies where LaMachine was used under
circumstances that go beyond this most common usage of LaMachine: using LaMachine on a secure server without internet
connection, and using LaMachine to provide a work environment for a group of researchers. These case studies illustrate
the advantages and limitations of LaMachine in practice.

@@

Most users of LaMachine are installing the software distribution in their own machine, either locally, via Docker or VM. However, we described to use cases in section \ref{sec:case} where we explore the possibilities under more constrained and demanding circumstances.
Such case studies contribute to thorough testing and running of LaMachine in less ideal circumstances such as nested VM constructions and offline usage.

@@

Development of LaMachine presently takes place in collaboration with the CLARIAH WP3 Virtual Research Environment (VRE)
project\footnote{\url{https://github.com/meertensinstituut/clariah-wp3-vre}}. The VRE project has higher ambitions in
accommodating the researcher and connectivity of data and services, and transcends also those of the CLARIN Language
Resource Switchboard~\cite{switchboard}. An important part of our future focus will therefore be on interoperability
with the higher-level tools emerging from the VRE efforts, but also with other parts of the CLARIN infrastructure;
single-sign on authentication being a notable example here.

>>>>

%\setlength\titlebox{5cm}
\setlist{nosep}
\setlist{noitemsep}
\setlength{\belowcaptionskip}{-10pt}
\setlength{\abovecaptionskip}{0pt}

@@

\section{Included Software}

@@

     Nijmegen:\footnote{Links: \url{https://languagemachines.github.io/timbl}, \url{https://languagemachines.github.io/mbt},

@@

 \item by the University of Groningen: \textbf{Alpino}\footnote{\url{http://www.let.rug.nl/vannoord/alp/Alpino/}} -- A dependency parser and tagger for Dutch.
 \item by the Vrije Universiteit Amsterdam:
     \textbf{KafNafParserPy}\footnote{\url{https://github.com/cltl/KafNafParserPy}} -- a python module to parse NAF
         files, another format for linguistic annotation; and
         \textbf{NafFoLiApy}\footnote{\url{https://github.com/cltl/NafFoLiApy}} -- converters between FoLiA and NAF.
 \item by Utrecht University: \textbf{T-scan}\footnote{\url{https://github.com/proycon/tscan}} -- a Dutch text analytics tool for readability prediction (initially developed at TiCC, Tilburg University, and in collaboration with Radboud University, Nijmegen)
 \item by the Meertens Instituut: \textbf{Python Course for the
     Humanities}\footnote{\url{http://www.karsdorp.io/python-course/}} -- interactive

@@

\begin{itemize}
    \item \textbf{Python}: Numpy, Scipy, Matplotlib, Scikit-Learn, IPython, Jupyter, ...
    \begin{itemize}
        \item \textbf{Jupyter Lab} The successor of the popular Jupyter Notebooks, offers notebooks, a web-based IDE, terminals. An ideal entry point to get started with LaMachine and all it contains!
        \item \textbf{PyTorch} - Deep-learning library for Python
        \item \textbf{NLTK} - Natural Language Toolkit for Python
        \item \textbf{Spacy} - Industrial-Strength NLP in Python
    \end{itemize}
    \item \textbf{R}, \textbf{Perl}, \textbf{Java}:
    \begin{itemize}
        \item \textbf{NextFlow} - A system and language for writing parallel and scalable pipelines in a portable manner.
        \item \textbf{Stanford CoreNLP} - Various types of linguistic enrichment
    \end{itemize}
    \item \textbf{Tesseract} - Open Source Optical Character Recognition (OCR)
    \item \textbf{Tensorflow} - Open-source machine learning framework
    \item \textbf{Kaldi} - Speech Recognition Framework (ASR)
    \item \textbf{Moses} - Statistical Machine Translation system
\end{itemize}

@@

\begin{figure}[htb] \begin{center}
\includegraphics[width=120.0mm]{screenshot_bootstrap.jpg}
\end{center}
\caption{\footnotesize{A screenshot of the bootstrap procedure}}
\label{fig:bootstrap}
\end{figure}

@@

\begin{figure}[htb] \begin{center}
\includegraphics[width=120.0mm]{screenshot_venv_activate.jpg}
\end{center}
\caption{\footnotesize{Activation of a local LaMachine environment on the command-line}}
\label{fig:venv}
\end{figure}

@@

\begin{figure}[htb] \begin{center}
\includegraphics[width=150.0mm]{screenshot_portal.jpg}
\end{center}
\caption{\footnotesize{A screenshot of the LaMachine portal, powered by Labirinto:
\url{https://github.com/proycon/labirinto}}}
\label{fig:portal}
\end{figure}

@@

\begin{figure}[htb] \begin{center}
\includegraphics[width=150.0mm]{screenshot_lab.jpg}
\end{center}
\caption{\footnotesize{A screenshot of the Jupyter Lab Environment}}
\label{fig:lab}
\end{figure}

@@

There are also CLARIN initiatives that provide not just a switchboard, but a more elaborate high-level virtual research
environment (VRE) with graphical interfaces to accommodate the researcher and allow execution and visualisation of
various NLP pipeline. The Dutch CLARIAH WP3 project\footnote{\url{https://github.com/meertensinstituut/clariah-wp3-vre}}
is such an initiative and LaMachine is currently developed in close collaboration with this project, in a role of technology supplier.

@@

LaMachine is most often used by a researcher or data scientist who installs LaMachine on his own computer (in a
local environment, Docker container or VM). Here we discuss two case studies where LaMachine was used under
circumstances that go beyond this most common usage of LaMachine: using LaMachine on a secure server without internet
connection, and using LaMachine to provide a work environment for a group of researchers. These case studies illustrate
the advantages and limitations of LaMachine in practice.

@@

Most users of LaMachine are installing the software distribution in their own machine, either locally, via Docker or VM.
However, we described two case studies in section~\ref{sec:case} where we explore the possibilities under more
constrained, complex, and demanding circumstances, which contributes to thorough testing and running of LaMachine.

@@

An important part of our future focus will be on interoperability with the higher-level tools emerging from
the CLARIAH WP3 VRE effort, but also with other parts of the CLARIN infrastructure; single-sign on authentication being
a notable example here.
